---
title: "Ferritin ID Thresholds"
format: html
---

```{r}
#| label: setup
FIGUREPATH <- "~/who_ferritin_comment/results/figures/"
BOOTPATH <- "~/who_ferritin_comment/results/boot/"
DATAPATH <- "~/who_ferritin_comment/data/"
MODELPATH <- "~/who_ferritin_comment/results/models/"
knots <- 3:25
response <- "Hemoglobin"
zoom <- T
boot_n <- 1000
```

```{r}
#| label: libraries
#| code-fold: true
#| code-summary: Libraries
#| message: false
#| warning: false

source("~/who_ferritin_comment/src/functions.R")
library(dplyr)
library(rms)
library(mcp)
library(boot)
library(scam)
```

## Load data

```{r}
# Load data
# Naming assumptions:
# Ferritin, Hemoglobin, sTfR
data <- readRDS(paste0(DATAPATH, "NHANES_variables_filtered.rds"))
```

## Test RCS model selection

(If plots are too small due to layout settings, you may change the chunk option `layout-ncol: 4` to 3, 2, or 1. You may also view these plots by accessing your FIGUREPATH, or by right-clicking here and selecting _Open image in new tab_.)

```{r}
#| label: test-rcs-knots
#| code-fold: true
#| code-summary: Visualize RCS knots
#| message: false
#| warning: false
#| fig-align: center
#| layout-ncol: 4

data_full <- data
data <- data %>% select(Ferritin, as.name(response))
# set datadist
dd <- datadist(data)
options(datadist = "dd")

investigate_RCS(data, response, knots, zoom, FIGUREPATH)
```

## Bootstrap knot selection

```{r}
#| label: boot-select-knot_n
#| code-fold: true
#| code-summary: Select optimum number of knots via bootstrapping and AIC
#| message: false
#| warning: false
#| fig-align: center

# The function boot_knot_selection is defined in functions.R
if (!file.exists(paste0(BOOTPATH, response, "_rcs_knots_", boot_n, "_bootobj.rds"))) {
    boot_knots <- boot(data = data, statistic = boot_knot_selection, R = boot_n, response = response, knots = knots)
    boot_knot_ci <- boot.ci(boot_knots, type = "perc") # BCa not computable on discrete data with this tight interval
    saveRDS(boot_knots, paste0(BOOTPATH, response, "_rcs_knots_", boot_n, "_bootobj.rds"))
    saveRDS(boot_knot_ci, paste0(BOOTPATH, response, "_rcs_knot_ci_", boot_n, "_boot_ci.rds"))
} else {
    boot_knots <- readRDS(paste0(BOOTPATH, response, "_rcs_knots_", boot_n, "_bootobj.rds"))
    boot_knot_ci <- readRDS(paste0(BOOTPATH, response, "_rcs_knot_ci_", boot_n, "_boot_ci.rds"))
    }

boot_df <- data.frame(t = boot_knots$t) %>% 
    mutate(within_ci = (t >= boot_knot_ci$percent[4] & t <= boot_knot_ci$percent[5]))

bootknot_ci_p <- ggplot(data = boot_df, aes(x = t, y = ..count../sum(..count..), fill = within_ci)) + 
    geom_bar() +
    theme_minimal() +
    labs(x = "Selected number of knots",
         y = "Frequency",
         fill = "Within CI")

ggsave(paste0(FIGUREPATH, response, "_rcs_boot_knot_distr.pdf"), bootknot_ci_p, width = 8, height = 6, device = cairo_pdf)

bootknot_ci_p

selected_knot_n <- boot_knots$t0

# Final RCS fit
rcs_fit <- ols( as.formula(sprintf("%s ~ rcs(Ferritin, %s)", response, selected_knot_n)), data = data, x = T, y = T)
saveRDS(rcs_fit, paste0(MODELPATH, "rcs_final.rds"))
```

## Bootstrap threshold

Now use selected knot number to bootstrap saddle estimates

```{r}
#| label: boot-find-saddle
#| code-fold: true
#| code-summary: Find estimates for saddle point via bootstrapping
#| message: false
#| warning: false
#| fig-align: center

# The function boot_saddle  is defined in functions.R
if (!file.exists(paste0(BOOTPATH, response, "_rcs_saddle_", boot_n, "_bootobj.rds"))) {
    boot_saddle <- boot(data = data, statistic = boot_saddle_estimate, R = boot_n, response = response, knot_n = selected_knot_n)
    invalid_indices <- which(boot_saddle$t == 0)
    boot_saddle_ci <- boot.ci(boot_saddle, type = "perc", exclude = invalid_indices)
    saveRDS(boot_saddle, paste0(BOOTPATH, response, "_rcs_saddle_", boot_n, "_bootobj.rds"))
    saveRDS(boot_saddle_ci, paste0(BOOTPATH, response, "_rcs_saddle_ci_", boot_n, "_boot_ci.rds"))
} else {
    boot_saddle <- readRDS(paste0(BOOTPATH, response, "_rcs_saddle_", boot_n, "_bootobj.rds"))
    boot_saddle_ci <- readRDS(paste0(BOOTPATH, response, "_rcs_saddle_ci_", boot_n, "_boot_ci.rds"))
    invalid_indices <- which(boot_saddle$t == 0)
    }

paste0("Bootstrapped saddle point estimate is computable in ", round((1 - length(invalid_indices)/length(boot_saddle$t)) * 100, 2), "% of iterations.")

rcs_final_p <- plot_rcs(data, rcs_fit, n_knot = selected_knot_n, zoom = T, FIGUREPATH = NULL) +
    annotate('rect', xmin = boot_saddle_ci$percent[4], xmax = boot_saddle_ci$percent[5], ymin = -Inf, ymax = Inf, alpha = 0.2, fill = 'black') +
    labs(subtitle = paste0("Saddle point: ", round(boot_saddle_ci$t0, 2), " μg/L [", round(boot_saddle_ci$percent[4], 2), ", ", round(boot_saddle_ci$percent[5], 2), "]"))

ggsave(paste0(FIGUREPATH, response, "_rcs_final.pdf"), rcs_final_p, width = 8, height = 6, device = cairo_pdf)

rcs_final_p

rcs_final_saddle_density_p <-
    ggplot(data = data.frame(t = boot_saddle$t), aes(x = t)) +
    geom_density() +
    theme_minimal() +
    labs(x = "Saddle point",
         y = "Density")

ggsave(paste0(FIGUREPATH, response, "_rcs_final_saddle_density.pdf"), rcs_final_saddle_density_p, width = 8, height = 6, device = cairo_pdf)

rcs_final_saddle_density_p
```


## Bayesian breakpoint analysis

We limit our data to here to Ferritin <= 100, otherwise the 2 breakpoint version will place the second breakpoint way beyond it. **This decision is a source of bias.**

### 1 breakpoint
```{r}
#| label: bayesian-bp1
#| code-fold: true
#| code-summary: Bayesian breakpoint analysis, 1 breakpoint
#| message: false
#| warning: false
#| fig-align: center

concat_data <- data %>% 
    filter(Ferritin <= 100)

if (!file.exists(paste0(MODELPATH, response, "_bayesian_1bp.rds"))) {
    # Specify model
    if (response == "Hemoglobin") {
        baye_1bp_spec <- list(Hemoglobin ~ Ferritin,
                              ~ 0 + Ferritin)
    }
    if (response == "sTfR") {
        baye_1bp_spec <- list(sTfR ~ Ferritin,
                              ~ 0 + Ferritin)
    }
    
    # Fit
    baye_1bp_fit <- mcp(baye_1bp_spec, 
                   data = concat_data,
                   adapt = 500,
                   iter = 500,
                   chains = 4,
                   cores = 4)
    
    saveRDS(baye_1bp_fit, paste0(MODELPATH, response, "_bayesian_1bp.rds"))
} else {
    baye_1bp_fit <- readRDS(paste0(MODELPATH, response, "_bayesian_1bp.rds"))
    }

summary(baye_1bp_fit)

# Plot
bay1_p <- plot(baye_1bp_fit) +
    theme_minimal() +
    coord_cartesian(xlim = c(0, 100))

posterior_check_p <- pp_check(baye_1bp_fit)

convergence_check_p <- plot_pars(baye_1bp_fit, regex_pars = "cp_")

ggsave(paste0(FIGUREPATH, response, "_bay1bp.pdf"), bay1_p, width = 8, height = 6, device = cairo_pdf)
ggsave(paste0(FIGUREPATH, response, "_bay1bp_ppcheck.pdf"), posterior_check_p, width = 8, height = 6, device = cairo_pdf)
ggsave(paste0(FIGUREPATH, response, "_bay1bp_convcheck.pdf"), convergence_check_p, width = 8, height = 6, device = cairo_pdf)

bay1_p
posterior_check_p
convergence_check_p
```

### 2 breakpoints
```{r}
#| label: bayesian-bp2
#| code-fold: true
#| code-summary: Bayesian breakpoint analysis, 2 breakpoints
#| message: false
#| warning: false
#| fig-align: center

if (!file.exists(paste0(MODELPATH, response, "_bayesian_2bp.rds"))) {
    # Specify model
    if (response == "Hemoglobin") {
        baye_2bp_spec <- list(Hemoglobin ~ Ferritin,
                              ~ 0 + Ferritin,
                              ~ 1 + Ferritin)
    }
    if (response == "sTfR") {
        baye_2bp_spec <- list(sTfR ~ Ferritin,
                              ~ 0 + Ferritin,
                              ~ 1 + Ferritin)
    }
    
    # Fit
    baye_2bp_fit <- mcp(baye_2bp_spec, 
                   data = concat_data,
                   adapt = 500,
                   iter = 500,
                   chains = 4,
                   cores = 4)
    
    saveRDS(baye_2bp_fit, paste0(MODELPATH, response, "_bayesian_2bp.rds"))
} else {
    baye_2bp_fit <- readRDS(paste0(MODELPATH, response, "_bayesian_2bp.rds"))
}

summary(baye_2bp_fit)

# Plot
bay2_p <- plot(baye_2bp_fit) +
    theme_minimal() +
    coord_cartesian(xlim = c(0, 100))

posterior_check_p <- pp_check(baye_2bp_fit)

convergence_check_p <- plot_pars(baye_2bp_fit, regex_pars = "cp_")

ggsave(paste0(FIGUREPATH, response, "_bay2bp.pdf"), bay2_p, width = 8, height = 6, device = cairo_pdf)
ggsave(paste0(FIGUREPATH, response, "_bay2bp.pdf"), posterior_check_p, width = 8, height = 6, device = cairo_pdf)
ggsave(paste0(FIGUREPATH, response, "_bay2bp.pdf"), convergence_check_p, width = 8, height = 6, device = cairo_pdf)

bay2_p
posterior_check_p
convergence_check_p
```

## Play with `scam`

```{r}
scamfit <- scam(Hemoglobin ~ s(Ferritin, k = 8, bs = "mpi"), data = concat_data)
plot(scamfit, residuals = T)
```

From the supplement of Galetti et al. (2021):

> To identify thresholds, we then used the method of finite differences to estimate the first derivatives of the fitted trend spline from the generalized additive modelling smoother 
> and we identified the periods in which the estimated rate of change between fitted y and x was statistically significant (i.e. statistically different from zero) by computing the
> uncertainty around the estimates from 10 000 posterior simulations. At the point on the x-axis where 95% of the posterior simulations are all different from zero for the first
> derivative, we could identify where the slope between fitted y and x began to be signiicantly positive (or negative).

The issue currently is that we are deliberately choosing to shape restrict the fit to a monotonically increasing curve $\rightarrow$ the derivative can never reach or cross zero, so we'd have to choose an arbitrary number "close to zero" instead, which will be difficult to justify.

```{r}
d1 <- derivative.scam(scamfit)
xx <- sort(concat_data$Ferritin, index = TRUE)
plot(xx$x, d1$d[xx$ix], type = "l")
```

